Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-06-08 hpmoon-asoco.tex
some other typos fixed

\documentclass[preprint]{elsarticle}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{moreverb,url,color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

@@

A similar approach used to solve this kind of problems is the {\em
  selective operator application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. On that paper we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other kind of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separating it in processors by objective \citep{Xiao03specialized}. \textcolor{blue}{For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector.} Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
\textcolor{blue}{Many real-world problems require high-dimensionality \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here.} The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. \textcolor{blue}{Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality}.

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier (from 1 to $P$). Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel}, all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\begin{figure}
\centering
\includegraphics[width=6cm]{hypervolume.jpg}
\caption{Hypervolume calculation. Areas from Pareto Front solutions (black points) to a reference point (white point) are added.}
\label{fig:hypervolume}
\end{figure}

@@

The previously described four approaches have been run with two
different chromosome lengths ($L$): 512 and 2048. Different number of
islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This
maximum number of islands has also been  used in a previous work in the
literature \cite{Martens13asynchronous}. The crossover and mutation
chosen, SBX and polynomial, have also been  used previously by other
authors in \cite{Durillo08masterslave}.  % SBX is first mentioned
                                % here. Has to explain - JJ

@@

\textcolor{blue}{The global population size is set to 1024. We have decided to use that value for two reasons: the first is that it is large enough to have 8 individuals per island by dividing that number when using 128 islands and small enough to have 128 individuals when using 8 islands. This is the minimum and maximum population size per island used by M{\''a}rtens et al. in his work \cite{Martens13asynchronous}.}

@@

Different metrics, previously explained, have been used
to calculate the quality of the obtained PFs in each configuration. \textcolor{blue}{As HV require a reference point to be
calculated (see Figure \ref{fig:hypervolume}), we have chosen the value (1,9) as none of the existing solutions on all fronts obtained during all executions are dominated by it. In other words, as we are minimizing in both objectives, this reference point is required to be numerically larger or equal in each objective, and strictly larger in at least one of them (dominated) for all solutions in the PFs we are going to measure. HV is then
normalized respecting to that point. Also, the optimal reference PF distribution used to calculate the IGD and Spread metrics has 1000 solutions \footnote{Optimal PFs are available at:   \url{http://www.tik.ee.ethz.ch/sop/download/supplementary/testproblems/}.} }.  
A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and Table
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained with D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$). We refer this equivalence with the word {\em Equiv-} in tables. For example, Equiv-D in an HV cell  means that the HV of A is the same than the value in D column, because the behaviour of this configuration is the same than executing D.

@@

%%%%%%FERGU: WARNING, THIS SECTION IS STILL SHIT
%ZDT1 
\textcolor{blue}{Figures \ref{fig:plot_zdt1_512} to \ref{fig:plot_zdt6_2048}, showing the final PFs of one arbitrary sample run of each combination. These figures allow to visualize the effect of the different chromosome sizes and number of islands in each different benchmark. The first one, ZDT1, has a convex Pareto-optimal front. Comparing results from $L=512$ (Figure \ref{fig:plot_zdt1_512}) and $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) it is apparent that the quality of the generated PFs depends on the chromosome size. Specifically, with $L=512$ it can be seen that the PFs quality decreases in Spread and HV with all configurations, but the SLA methods behave worse than $B$, as it also was shown in average in Table \ref{tab:results512}. Nevertheless, if we focus on the plots obtained with $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) this difference is less clear. It can be seen in Figure \ref{fig:plot_zdt1_2048} that solutions obtained by the SLA methods are more dispersed than the baseline as $P$ grows. However, with $A$, the area covered by the PF is still larger.}

%ZDT2
\textcolor{blue}{ZDT2 has a non-convex Pareto front. From Figures \cite{fig:plot_zdt2_512} and \cite{fig:plot_zdt2_2048} is clear that, as in the previous problem ZDT1, there is a decrease of quality for all configurations as the size of $P$ increases.  Nevertheless, for this benchmark, there is not a clear decrease in the PF quality when $L$ goes from 512 to 2048.}

%ZDT3
\textcolor{blue}{ZDT3 Pareto-optimal front is formed by non-contiguous convex parts, and plots for both chromosome sizes (Figures \ref{fig:plot_zdt3_512} and \ref{fig:plot_zdt3_2048}) show solutions gathered in similar regions in all configurations. However, with $L=512$, the rightmost part of the obtaninde non-dominated solutions is not obtained in the $A$ alternative with $P=8$  and $P=16$ (being equivalent to alternatives $O$ and $D$). This explains why with this chromosome length, the SLA methods not improve the results obtained by $B$. This problem does not appear when $L=2048$. Although the differences in the Spread metric are not significant for the different alternatives for this benchmark, (as shown in Table \ref{tab:results2048}), the HV and IGD metrics are the ones that can be used to show the effectiveness of our approach in this type of problem with non-continuous PFs.}

\textcolor{blue}{In ZDT6, the Pareto-optimal solutions are not only non-uniformly distributed along the PF, but their density also increases as we move further away from it. That is the reason why some PFs obtained have an extremely low number of solutions in some configurations, and also, they are aggregated toward the areas with more density. Moreover, all configurations, including $B$, obtained PFs relatively far away from the optimal PF, as opposed as in previous problems. Spread is very similar for all methods $B$, $D$, $O$, $A$ across different $L$ sizes. Indeed, there is not significant difference between methods, with an exception: Spread using $A$ with $L=512$ is significantly worse. As in previous approaches it is visually clear that increasing the number of islands, and therefore, reducing the number of individuals per island, imply a decrease in quality of the HV.}

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. \textcolor{blue}{Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.}
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending \textcolor{blue}{on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on results.
% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ Pablo: extending this section

>>>>

\documentclass[preprint]{elsarticle}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{moreverb,url,color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
%\usepackage{refcheck}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

@@

A similar approach used to solve this kind of problems is the {\em
  selective operator application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
\textcolor{blue}{Another kind of approaches has also been explored}. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, the authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separating it in processors by objective \citep{Xiao03specialized}. \textcolor{blue}{For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector.} Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
\textcolor{blue}{Many real-world problems require high-dimensionality \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here.} The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. \textcolor{blue}{Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality. Others authors, such as Cao et al. \cite{CaoZLL17} separate the decision variables into several subpopulations after performing an interdependence analysis, but using a fixed number of processors.}

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier (from 1 to $P$). \textcolor{blue}{Note that, unlike the other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel}}, all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\begin{figure}
\centering
\includegraphics[width=6cm]{hypervolume.jpg}
\caption{Hypervolume calculation. \textcolor{blue}{Areas from Pareto Front solutions (black points) to a reference point (white point) are used to calculate the value of the hypervolume (all the gray area)}.}
\label{fig:hypervolume}
\end{figure}

@@

The previously described four approaches have been run with two
different chromosome lengths ($L$): 512 and 2048. Different number of
islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This
maximum number of islands has also been  used in a previous work in the
literature \cite{Martens13asynchronous}. The crossover and mutation
chosen, SBX and polynomial, have also been  used previously by other
authors in \cite{Durillo08masterslave,CaoZLL17}.  % SBX is first mentioned
                                % here. Has to explain - JJ

@@

\textcolor{blue}{The global population size is set to 1024. We have decided to use that value for two reasons: the first is that it is large enough to have 8 individuals per island by dividing that number when using 128 islands and small enough to have 128 individuals when using 8 islands. This is the minimum and maximum population size per island used by M{\"a}rtens et al. in his work \cite{Martens13asynchronous}.}

@@

Different metrics, previously explained, have been used
to calculate the quality of the obtained PFs in each configuration. \textcolor{blue}{As HV require a reference point to be
calculated (see Figure \ref{fig:hypervolume}), we have chosen the value (1,9) as none of the existing solutions on all fronts obtained during all executions are dominated by it. In other words, as we are minimizing in both objectives, this reference point is required to be numerically larger or equal in each objective, and strictly larger in at least one of them (dominated) for all solutions in the PFs we are going to measure. HV is then
normalized respecting to that point. Also, the optimal reference PF distribution used to calculate the IGD and Spread metrics has 1000 solutions \footnote{Optimal PFs are available at:   \url{http://www.tik.ee.ethz.ch/sop/download/supplementary/testproblems/}.} }.  
A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and Table
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained by D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$). We refer this equivalence with the word {\em Equiv-} in tables. For example, Equiv-D in an HV cell  means that the HV of A is the same than the value in D column, because the behaviour of this configuration is the same than executing D.

@@

%%%%%%FERGU: WARNING, THIS SECTION IS STILL SHIT
%ZDT1 
\textcolor{blue}{Figures \ref{fig:plot_zdt1_512} to \ref{fig:plot_zdt6_2048}, showing the final PFs of one arbitrary sample run of each combination. These figures allow to visualize the effect of the different chromosome sizes and number of islands in each different benchmark. The first one, ZDT1, has a convex Pareto-optimal front. Comparing results from $L=512$ (Figure \ref{fig:plot_zdt1_512}) and $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) it is apparent that the quality of the generated PFs depends on the chromosome size. Specifically, with $L=512$ it can be seen that the PFs quality decreases in Spread and HV with all configurations, but the SLA methods behave worse than $B$, as it also was shown on average in Table \ref{tab:results512}. Nevertheless, if we focus on the plots obtained with $L=2048$ (Figure \ref{fig:plot_zdt1_2048}) this difference is less clear. It can be seen in Figure \ref{fig:plot_zdt1_2048} that solutions obtained by the SLA methods are more dispersed than the baseline as $P$ grows. However, with $A$, the area covered by the PF is still larger.}

%ZDT2
\textcolor{blue}{ZDT2 has a non-convex Pareto front. From Figures \ref{fig:plot_zdt2_512} and \ref{fig:plot_zdt2_2048} is clear that, as in the previous problem ZDT1, there is a decrease of quality for all configurations as the size of $P$ increases.  Nevertheless, for this benchmark, there is not a clear decrease in the PF quality when $L$ goes from 512 to 2048.}

%ZDT3
\textcolor{blue}{ZDT3 Pareto-optimal front is formed by non-contiguous convex parts, and plots for both chromosome sizes (Figures \ref{fig:plot_zdt3_512} and \ref{fig:plot_zdt3_2048}) show solutions gathered in similar regions in all configurations. However, with $L=512$, the rightmost part of the obtaninde non-dominated solutions is not obtained in the $A$ alternative with $P=8$  and $P=16$ (being equivalent to alternatives $O$ and $D$). This explains why with this chromosome length, the SLA methods do not improve the results obtained by $B$. This problem does not appear when $L=2048$. Although the differences in the Spread metric are not significant for the different alternatives for this benchmark, (as shown in Table \ref{tab:results2048}), the HV and IGD metrics are the ones that can be used to show the effectiveness of our approach in this type of problem with non-continuous PFs.}

\textcolor{blue}{In ZDT6, the Pareto-optimal solutions are not only non-uniformly distributed along the PF, but their density also increases as we move further away from it. That is the reason why some PFs obtained have an extremely low number of solutions in some configurations, and also, they are aggregated toward the areas with more density. Moreover, all configurations, including $B$, obtained PFs relatively far away from the optimal PF, as opposed as in previous problems. Spread is very similar for all methods $B$, $D$, $O$, $A$ across different $L$ sizes. Indeed, there is not a significant difference between methods, with an exception: Spread using $A$ with $L=512$ is significantly worse. As in previous approaches it is visually clear that increasing the number of islands, and therefore, reducing the number of individuals per island, imply a decrease in quality of the HV.}

@@

Obtained results also show that when increasing the number of islands, the
automatic overlapping method ($A$) significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in automatic overlapping
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. \textcolor{blue}{Since our method divides the individuals into blocks of divisible size, aggregation methods after migration, such as the ones used in \cite{Kimovski15Parallel} or \cite{Dorronsoro13superlinear} can be also tested. But as we are dealing with complete individuals in each island, the possibility of using a real number, such as 0.3 (30\% overlapping) instead an integer value opens future experimental setups.}
For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending \textcolor{blue}{on population size, number of islands, or other values, or using some statistical parameter setting method, such as Taguchi's \cite{Keratmatpour2018Taguchi}}. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on the results.
% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ Pablo: extending this section
